---
title: "Conditioned Probabilities in Natural Hazards"
author: "Nora Krebs and Rebecca Amberger"
date: "27 November 2018"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task of the week

This week our task was to calculate the conditioned probability of a new problem, using Bayes' law again and to present the results in a Shiny app. 

The new conditioned probability problem is the following: 

"Magician Michelle says that she can predict large megathrust earthquakes. well, at least 95% of the time. Assume that these earthquakes occur with a probability of $10^{-5}$. How likely will Michelle cause a false alarm with her next prediction?"

## Analyzing the given problem

The problem can be solved with Bayes' law:

$$P(A|B)=\frac{P(B|A)*P(A)}{P(B)}$$
the given variables are:

E = the earthquake occurres\    
EC = the earthquake does not occurre\    
M = Michelle is right\    
MC = Michelle is wrong\   

```{r}
#probability of earthquake occurrence:
  E <- 0.00001
  E * 100 #in %
#probability of no earthquake occurrence:
  EC <- 1-E
  EC * 100 #in %
#probability of Michelle being right:
  M <- 0.95
  M * 100 #in %
#probability of Michelle being wrong:
  MC <- 1-M
  MC * 100 #in %
```

If we pair the tuples in a tree diagram, we would get the following probabilities:

```{r}
#Earthquake appears and Michelle is right
E_M <- E * M
E_M * 100 #in %

#Earthquake appears and Michelle is wrong
E_MC <- E * MC
E_MC * 100 #in %

#Earthquake does not appear and Michelle is right
EC_M <- EC * M
EC_M * 100 #in %
  
#Earthquake does not appear and Michelle is wrong
EC_MC <- EC * MC
EC_MC * 100 #in %

```

##Solution of the problem with Bayes' theorem

In the specific case we are looking for the probability that Michelle is wrong about her earthquake prediction (she forecasts an earthquake, but no earthquake occurres). Therefore we are most interested in the cases of $E_MC$ and $EC_MC$

If we apply Bayes' theorem, the formula can be written in the following way:

$$P(EC|MC)=\frac{P(MC|EC)*P(EC)}{P(MC)}$$

We can then solve the problem in the following way:

$$P(EC|MC)=\frac{P(MC|EC)*P(EC)}{\sum\limits_{EC} P(MC|EC)*P(EC)}$$

```{r}
#conditioned probability that the earthquake does not appear, given that Michelle says that there will be an earthquake
EC_con_MC <- EC_MC/(EC_MC + E_MC)
EC_con_MC * 100 #in %
```

###Answer to the question

Michelle will cause a false alarm with a probability of almost 100%, given that the earthquake does not occure.

##Discussion of the problem

Although Michelle is good at predicting earthquakes, by giving a right prediction in 95 % of the cases, the event of a large megathrust earthquake is so unlikely (only 0.001 % ) that she barely even has the chance to give a right prediction and ends up givig fals warnings almost 100 % of the time.

It would be interesting now to see how the prediction rate of a person and the likeliness of an event to happen are influencing eachother. Therefore we have developed a App in R with the Shiny Package.

##Shiny App for Conditioned probabilites in Natural Hazards

```{r}
# Starts a shiny app:
library(shiny)

# Defines user interface for application that draws a plot:
ui <- fluidPage(
  
  # Application title
  headerPanel("Conditioned Probability in Natural Hazards"),
  # Authors and title
  titlePanel("Interactive Application by Nora Krebs and Rebecca Amberger"),
  #Date
  titlePanel("created on: 26/11/2018"),
  
 # Layout of the Sidebar with input(prediction rate) and output (wrong forecast):
  sidebarLayout(
    #sidebar panel for inpu:
    sidebarPanel(
      #Input: slider to set a certain prediciton rate and earthquake probability
      sliderInput(inputId = "Prediction_Rate",
                  label = "Prediction Rate in %:",
                  min = 0,
                  max = 100,
                  value = 95),
      
      sliderInput(inputId = "Earthquake_Probability",
                  label = "Earthquake Probability in %:",
                  min = 0,
                  max = 1,
                  value = 0.00001)
    ),

    # Main panel shows a plot that displays the output
    mainPanel(
      # Output: Plot of the Wrong Forecast versus Event Occurrence:
      plotOutput(outputId = "Wrong_Forecast")
    )
  )
)
```

```{r}
# Define requirements which are needed by server to draw a plot
server <- function(input, output) {
  
  #making the plot interactive, so it is re-drawn after the input has been changed 
  #(by using the 'renderPlot'-function):
  output$Wrong_Forecast <- renderPlot({
    
    #creating a vector that defines the input for x: 'Event Occurrence'
     x <- c(0.0000001, 0.0000001, 0.000001, 0.00001, 0.0001, 0.001, 0.01, 0.1, 1);
     
     #Given Baye´s theorem, we want to calculate the probability of a wrong forecast, given that a hazardous 
     #event occurres at a certain probability, given in the vector x
     
     y <- (((1-input$Prediction_Rate * 1-x) / ((1-input$Prediction_Rate * 1-x) + (1-input$Prediction_Rate * x)))* 100)
       
    #creates a graph with axes, titles, etc., while "type=0" indicates that points are not plotted yet
    #and "log=x" indicates that the 'Event Occurrence' is plotted on a logarithmic scale and indicates all
    #further settings of the plot (caption, axis labels, magnification of labels/caption)
    plot(x,y, log="x", type= "n", main = "Probability of a wrong Forecast with changing Event probabilities, 
         for a known Prediciton rate",
         xlab = "Event Occurrence [%]", ylab = "Wrong Forecast [%]",
         cex.main = 1.4, cex.lab = 1.2)
    #creates a line chart where points are overplotted by lines
    lines(x,y,type= "o")
  })
}

# Run the application 
shinyApp(ui = ui, server = server)
```

